---
title: Explore
author: Andy Choens
output: html_document
---

```{r setup, include = FALSE}
pacman::p_load(knitr, rio, tidyverse)
baby_names <- import("data/baby_names.rds")
census_regions <- import("data/census_regions.rds")

## Placeholder for joins

```



# Data Exploration

## Do we have all the states/years?

- State: All
- Year: Min/Max
- Sex: All
- Names: How many unique

```{r echo=FALSE}
baby_names |> 
  group_by(state_code) |>
  summarize(
    first_yob = min(yob),
    last_yob = max(yob),
    unique_names = length(unique(name))
  ) |> 
  kable()
```


```{r problems, echo=FALSE, results = 'asis'}
problems <- 
  baby_names |> 
  group_by(state_code) |>
  summarize(
    first_yob = min(yob),
    last_yob = max(yob),
    unique_names = length(unique(name))
  ) %>%
  filter(first_yob != 1910 | last_yob != 2021)

if (nrow(problems) > 0) {
  cat("## Problems:")
  cat("\n\n")
  cat("The following rows have problems:")
  cat("\n\n")
  problems |> kable()
}
```

## The Last Homer

When was the last year there was someone named Homer born in Oregon?

```{r}
baby_names |>
  filter(state_code == "OR", name == "Homer") |>
  arrange(desc(yob)) |>
  top_n(2) |>
  kable()
```

## New York State Popular Names (2021)

- State: New York
- Year: 2021
- Sex: Male
- Names: 10 most popular

```{r echo=FALSE}
baby_names |>
  filter(state_code == "NY", yob == 2021,  sex == "M") |>
  arrange(desc(n_occurrences)) |>
  top_n(10) |>
  kable()
```

Make a bar plot of the top 25 most popular names in New York in 2021 and show the differences in the counts!

```{r}
baby_names |>
  filter(state_code == "NY", yob == 2021,  sex == "M") |>
  arrange(desc(n_occurrences)) |>
  top_n(25) |>
  ggplot(aes(x = n_occurrences, y = name)) +
  geom_col() +
  labs(title = "Most Popular Male Names in New York, 2021")+
  xlab("Count") +
  ylab("Name")
```



# Gender Changes

## Charlie - What a guy?

- Charlie Brown is one of the most famous Charlie's of all time.
- Traditionally male name.
- What percent of Charlies born in the US in 2021 were female?
- Name: Charlie
- Year: 2021
- Sex: What percent is female?

```{r}
baby_names |>
  filter(yob == 2021, name == "Charlie") |>
  group_by(sex) |>
  summarize(n = sum(n_occurrences)) |>
  pivot_wider(names_from = sex, values_from = n) |>
  mutate(Total = F+M) %>%
  select("Female" = F, "Male" = M, Total = Total) |>
  mutate("Percent Female" = round(100*Female/Total,1))
```

Can we plot the change in Charlie over time?

```{r}

```

## Names that went from Male to Female

- What names were 90% or more male prior to 1950 which have become more than 50% female in the past 10 years?
- Year: 



# Exploring Regional Differences

We will do this Friday. If you have not yet done so, PLEASE PLEASE PLEASE review the reading.

## Northeast Popular Names (2021)

- Region: Northeast
- Year: 2021
- Sex: Male
- Names: 10 most popular

```{r}

```



# Challenge!

We'll come up with some challenges together to further explore this data on Friday.